SUPPORTING INFORMATION:\\ Headgroup Structure and Cation Binding in Phosphatidylserine Lipid Bilayers

### 2019-10-17 Manuscript/manuscriptPSsuppl.tex

\subsection{MacRog}
\noindent
{\it POPC bilayers.} 
% \todo{For example this data set has pretty good description of simulation details in the
% reference. I am not sure how much we should repeat these in here.}
The POPC bilayer consisting of a total of 128 lipids was constructed from single lipid structure taken 
from Ref.~\citenum{kulig15c}. The bilayer was hydrated by a total of 5120 water 
molecules, corresponding to 40 water molecules per lipid.
The MacRog lipid parameters \cite{kulig15b}, obtained from Ref.~\citenum{kulig15c}, were used,
along with the TIP3P water model~\cite{jorgensen83}. The simulation parameters were identical to those used for 
POPC/POPC mixtures with the MacRog force field, see below. The simulation was run for 200~ns using Gromacs 2016.3~\cite{abraham2015gromacs}, 
out of which 150~ns were used for analyses. The simulation data and related files are available from Ref. \citenum{macrogPOPC298K}.

\noindent
{\it POPS bilayers with Na$^+$ counterions.}
%The data in Table \ref{PSsystems} and Refs \citenum{macrogPOPS298Kcorrect,macrogPOPS298KwithK}
%\todo{T. Piggot and M. Javanainen, please write simulation details} \\
%\hl{Piggot: please write about system with Na$^+$ counterions}.
Force field parameters for POPS simulations with the MacRog OPLS-AA force field~\cite{maciejewski14,kulig15b,rog16}
were constructed manually using the PS parameters provided in ref.~\citenum{rog16} as a guide. The original parameters were
not used due to an incorrect arrangement of the {\it sn}-1 and {\it sn}-2 tails (i.e. the lipid was OPPS and not POPS).
The corrected parameters are available from Ref.~\citenum{macrogPOPS298KwithK}.
The starting structure was the same as that used in the CHARMM36 POPS simulations (with Na$^+$ ions) and was converted into
the correct format using a script to adjust the atom order. Water was treated using the TIP3P model~\cite{jorgensen83}.
Simulation parameters were chosen to closely mimic those used in the original force field publications~\cite{maciejewski14,kulig15b,rog16}:
van der Waals interactions were truncated at 1.0~nm with a dispersion correction applied to the energy and pressure;
Long-range Coulombic interactions were treated with PME beyond a system-optimized cut-off.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations,
including use of the Verlet cut-off scheme~\cite{Pall13}, apart from the use of LINCS to constrain all bonds during these simulations.
Simulations with the group cut-off scheme gave slightly lower area per molecule ($\sim$~50\AA$^2$) than
the simulations with Verlet cut-off ($\sim$~52\AA$^2$).

@@

\noindent
{\it DOPS and POPS bilayers.} 
%The data in Table \ref{PSsystems} and Refs \citenum{bergerPOPS298K,bergerDOPS303K}
%\todo{T. Piggot, please write simulation details} \\
Starting structures for DOPS and POPS simulations performed with a Berger-based PS force field
were taken from those produced for the CHARMM36 simulations. A script was used to re-order the
atoms and to remove any extraneous hydrogen atoms. Force field parameters were taken from those
published by Mukhopadhyay et al.~\cite{mukhopadhyay04}
% [REF: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1303994/]
and the SPC water model was used~\cite{berendsen81}. The simulation cut-offs closely matched
those of Mukhopadhyay et al.: van der Waals interactions were truncated at 1.0~nm;
Coulombic interactions were truncated at 1.0~nm with PME applied to account for the interactions beyond the cut-off.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations,
including the use of the Verlet cut-off scheme, apart from all bonds were constrained using LINCS in these simulations.

@@

\noindent
{\it DOPS and POPS bilayers.} 
%The data in Table \ref{PSsystems} and Refs. \citenum{ckp1POPS303K,ckp2POPS303K,ckp1DOPS303K,ckp2DOPS303K}
%\todo{T. Piggot, please write simulation details} \\
Starting structures for GROMOS-CKP and GROMOS-CKPM simulations were taken from those produced for the CHARMM36 simulations.
A script was used to re-order the atoms and to remove any extraneous hydrogen atoms. The force fields parameters for DOPS and POPS
were as discussed above. The SPC water model was used in these simulations~\cite{berendsen81}. Simulation cut-offs were chosen
to closely match those used in the original GROMOS-CKP force field validation: van der Waals interactions were truncated at 1.4~nm
with a dispersion correction applied to the energy and pressure; Coulombic interactions were also truncated at 1.4~nm with PME used
to treat the long-range interactions. All other simulation settings were identical to those described above for the CHARMM36 DOPS
and POPS simulations, including the use of the Verlet cut-off scheme, apart from all bonds were constrained using LINCS in these simulations.

@@

\begin{figure}[]
  \centering
  \includegraphics[width=17.0cm]{../Figs/OP_CHARMM_CaCl_POPC_NBFix.pdf}
  \caption{\label{OP_CHARMM_CaCl_POPC_NBFix}
    Headgroup order parameters from CHARMM36 simulations of POPC, where the NBfix term was employed employed for sodium \cite{venable13} ({\it left})
    and calcium \cite{kim16} ({\it right}) compared with the experimental data \cite{akutsu81,altenbach84}
    and simulations without NBfix for the calcium.
    Simulation files without ions are available at Ref. \citenum{nencineCHARMM36popcT310K},
    with the NBfix term in sodium at Ref.~\citenum{CHARMM36withCaClNBfix}, 
    with the NBfix term in calcium at Ref.~\citenum{CHARMM36withCaClNBfix} and
    without the NBfix in calcium at Ref.~\citenum{charmmPOPC450mMCaClfiles}.
  }
\end{figure}

>>>>

\subsection{MacRog}
\noindent
{\it POPC bilayers.} 
% \todo{For example this data set has pretty good description of simulation details in the
% reference. I am not sure how much we should repeat these in here.}
The POPC bilayer consisting of a total of 128 lipids was constructed from single lipid structure taken 
from Ref.~\citenum{kulig15c}. The bilayer was hydrated by a total of 5120 water 
molecules, corresponding to 40 water molecules per lipid.
The MacRog lipid parameters \cite{kulig15b}, obtained from Ref.~\citenum{kulig15c}, were used,
along with the TIP3P water model~\cite{jorgensen83}. The simulation parameters were identical to those used for 
POPC/POPS mixtures with the MacRog force field, see below. The simulation was run for 200~ns using Gromacs 2016.3~\cite{abraham2015gromacs}, 
out of which 150~ns were used for analyses. The simulation data and related files are available from Ref. \citenum{macrogPOPC298K}.

\noindent
{\it POPS bilayers with Na$^+$ counterions.}
%The data in Table \ref{PSsystems} and Refs \citenum{macrogPOPS298Kcorrect,macrogPOPS298KwithK}
%\todo{T. Piggot and M. Javanainen, please write simulation details} \\
%\hl{Piggot: please write about system with Na$^+$ counterions}.
Force field parameters for POPS simulations with the MacRog OPLS-AA force field~\cite{maciejewski14,kulig15b,rog16}
were constructed manually using the PS parameters provided in ref.~\citenum{rog16} as a guide. The original parameters were
not used due to an incorrect arrangement of the {\it sn}-1 and {\it sn}-2 tails (i.e. the lipid was OPPS and not POPS).
The corrected parameters are available from Ref.~\citenum{macrogPOPS298KwithK}.
The starting structure was the same as that used in the CHARMM36 POPS simulations (with Na$^+$ ions) and was converted into
the correct format using a script to adjust the atom order. Water was treated using the TIP3P model~\cite{jorgensen83}.
Simulation parameters were chosen to closely mimic those used in the original force field publications~\cite{maciejewski14,kulig15b,rog16}:
van der Waals interactions were truncated at 1.0~nm with a dispersion correction applied to the energy and pressure;
Long-range Coulombic interactions were treated with PME beyond a system-optimized cut-off.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations,
including use of the Verlet cut-off scheme~\cite{Pall13}, except that here LINCS was used to constrain all bonds.
Simulations with the group cut-off scheme gave slightly lower area per molecule ($\sim$~50\AA$^2$) than
the simulations with Verlet cut-off ($\sim$~52\AA$^2$).

@@

\noindent
{\it DOPS and POPS bilayers.} 
%The data in Table \ref{PSsystems} and Refs \citenum{bergerPOPS298K,bergerDOPS303K}
%\todo{T. Piggot, please write simulation details} \\
Starting structures for DOPS and POPS simulations performed with a Berger-based PS force field
were taken from those produced for the CHARMM36 simulations. A script was used to re-order the
atoms and to remove any extraneous hydrogen atoms. Force field parameters were taken from those
published by Mukhopadhyay et al.~\cite{mukhopadhyay04}
% [REF: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1303994/]
and the SPC water model was used~\cite{berendsen81}. The simulation cut-offs closely matched
those of Mukhopadhyay et al.: van der Waals interactions were truncated at 1.0~nm;
Coulombic interactions were truncated at 1.0~nm with PME applied to account for the interactions beyond the cut-off.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations,
including the use of the Verlet cut-off scheme, except that here LINCS was used to constrain all bonds.

@@

\noindent
{\it DOPS and POPS bilayers.} 
%The data in Table \ref{PSsystems} and Refs. \citenum{ckp1POPS303K,ckp2POPS303K,ckp1DOPS303K,ckp2DOPS303K}
%\todo{T. Piggot, please write simulation details} \\
Starting structures for GROMOS-CKP and GROMOS-CKPM simulations were taken from those produced for the CHARMM36 simulations.
A script was used to re-order the atoms and to remove any extraneous hydrogen atoms. The force fields parameters for DOPS and POPS
were as discussed above. The SPC water model was used in these simulations~\cite{berendsen81}. Simulation cut-offs were chosen
to closely match those used in the original GROMOS-CKP force field validation: van der Waals interactions were truncated at 1.4~nm
with a dispersion correction applied to the energy and pressure; Coulombic interactions were also truncated at 1.4~nm with PME used
to treat the long-range interactions. All other simulation settings were identical to those described above for the CHARMM36 DOPS
and POPS simulations, including the use of the Verlet cut-off scheme, except that here LINCS was used to constrain all bonds.

@@

\begin{figure}[]
  \centering
  \includegraphics[width=17.0cm]{../Figs/OP_CHARMM_CaCl_POPC_NBFix.pdf}
  \caption{\label{OP_CHARMM_CaCl_POPC_NBFix}
    Headgroup order parameters from CHARMM36 simulations of POPC, where the NBfix term was employed employed for sodium \cite{venable13} ({\it left})
    and calcium \cite{kim16} ({\it right}) compared with the experimental data \cite{akutsu81,altenbach84}
    and simulations without NBfix for the calcium.
    Simulation files without ions are available at Ref. \citenum{nencineCHARMM36popcT310K},
    with the NBfix term in sodium at Ref.~\citenum{CHARMM36withNaClNBfix}, 
    with the NBfix term in calcium at Ref.~\citenum{CHARMM36withCaClNBfix} and
    without the NBfix in calcium at Ref.~\citenum{charmmPOPC450mMCaClfiles}.
  }
\end{figure}
